---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "60%",
  dpi = 300,
  fig.width = 5,
  fig.height = 5,
  fig.align = "center",
  out.extra="style='text-align: center; display: block; margin: auto;'"
)
```

# `tna`: An R package for Transition Network Analysis

<!-- badges: start -->
[![R-CMD-check](https://github.com/sonsoleslp/tna/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sonsoleslp/tna/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/sonsoleslp/tna/graph/badge.svg)](https://app.codecov.io/gh/sonsoleslp/tna)
[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)

<!-- badges: end -->

An R package for the analysis of relational dynamics through 
Transition Network Analysis (TNA). TNA provides tools for building TNA models, 
plotting transition networks, calculating centrality measures, and identifying 
dominant events and patterns. TNA statistical techniques (e.g., bootstrapping 
and permutation tests) ensure the reliability of observed insights and confirm
that identified dynamics are meaningful. See (Saqr et al., 2024) 
<doi:10.48550/arXiv.2411.15486> for more details on TNA.

## Installation

You can install `tna` from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("sonsoleslp/tna")
```

## Example

Load the library
```{r}
library("tna")
```

Example data
```{r}
data("group_regulation", package = "tna")
```

Build a Markov model
```{r}
tna_model <- tna(group_regulation)
```

```{r, eval = F}
summary(tna_model)
```
```{r, echo=FALSE}
summary(tna_model) |>
    gt::gt() |>
    gt::fmt_number(decimals  = 2) |>
    gt::as_raw_html()
```


Plot the transition network
```{r tnaplot}
plot(tna_model)
```

Calculate the centrality measures
```{r}
cent <- centralities(tna_model)
```

```{r, echo=FALSE}
cent |>
    gt::gt() |>
    gt::tab_options(table.font.size = 8) |>
    gt::as_raw_html()
```

Plot the centrality measures
```{r centralitiesplot, fig.width=12, fig.height=8, out.width="90%"}
plot(cent, ncol = 3)
```

Estimate centrality stability
```{r}
estimate_centrality_stability(tna_model)
```


Identify and plot communities
```{r}
coms <- communities(tna_model)
plot(coms)
```

Find and plot cliques
```{r, out.width="50%"}
cqs <- cliques(tna_model, threshold = 0.2)
plot(cqs)
```

Compare high achievers (first 1000) with low achievers (last 1000)
```{r, eval=FALSE}
tna_model_start_high <- tna(group_regulation[1:1000, ])
tna_model_start_low <- tna(group_regulation[1001:2000, ])
comparison <- permutation_test(tna_model_start_high, tna_model_start_low,
                               measures = c("InStrength"))
plot(comparison)
```
 
::: {layout-ncol=2}
```{r, echo=FALSE}
tna_model_start_high <- tna(group_regulation[1:1000, ])
tna_model_start_low <- tna(group_regulation[1001:2000, ])
comparison <- permutation_test(tna_model_start_high, tna_model_start_low,
                               measures = c("InStrength"))
plot(comparison)
```

```{r, echo = FALSE}
comparison$centralities$stats |> gt::gt() |> 
    gt::as_raw_html()
```
:::


