---
title: "Getting started with the tna package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with the tna package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.height = 4,
  dev = "jpeg",
  dpi = 100,
  out.width = "100%",
  comment = "#>"
)
suppressPackageStartupMessages({
  library("tna")
  library("tibble")
  library("dplyr")
  library("gt")
})
```

This vignette showcases some basic usage of the `tna` package. For more tutorials, please visit [the package website](https://sonsoles.me/tna/).

First we load the package that we will use for this example.

```{r, eval = FALSE}
library("tna")
library("tibble")
library("dplyr")
library("gt")
```

We also load the `group_regulation` data available in the package (see `?group_regulation` for further information)
```{r}
data("group_regulation", package = "tna")
```

We build a TNA model using this data with the `tna()` function .
```{r}
tna_model <- tna(group_regulation)
```

To visualize the model, we can use the standard `plot()` function.
```{r, fig.width=5, fig.height=5}
plot(
  tna_model, cut = 0.2, minimum = 0.05, 
  edge.label.position = 0.8, edge.label.cex = 0.7
)
```

The initial state probabilities are
```{r}
data.frame(`Initial prob.` = tna_model$inits, check.names = FALSE) |>
  rownames_to_column("Action") |>
  arrange(desc(`Initial prob.`)) |>
  gt() |>
  fmt_percent()
```

and the transitions probabilities are
```{r}
tna_model$weights |>
  data.frame() |>
  rownames_to_column("From\\To") |>
  gt() |>
  fmt_percent()
```

The function `centralities()` can be used to compute various centrality measures (see `?centralities` for more information).
These measures can also be visualized with the `plot()` function.
```{r}
centrality_measures <- c("BetweennessRSP", "Closeness", "InStrength", "OutStrength")
cents_withoutloops <- centralities(
  tna_model,
  measures = centrality_measures,
  loops = FALSE,
  normalize = TRUE
)
plot(cents_withoutloops, ncol = 2, model = tna_model)
```
