---
title: "How to convert time series data into sequence data for `tna`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to prepare data for tna}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

TNA is intended for analysis of ordered sequential data (i.e., sequences). However, it supports the converstion from time series data (i.e., ordered continous data) to the format necessary for TNA.


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.height = 4,
  out.width = "100%",
  message = FALSE,
  fig.show = TRUE,
  results = "hide",  
  comment = "#>"
)
suppressPackageStartupMessages({
  library("tna")
  library("tibble")
  library("dplyr")
  library("tidyr")
})
```

To showcase this feature, we will use the `EuStockMarkets`, which is included by default in R as a multivariate time series object representing the daily closing prices of major European stock indices (DAX, SMI, CAC, and FTSE) from 1991 to 1998.

We begin by converting the EuStockMarkets object into a long data frame, retaining the timepoints as the first column, the stock as the second column, and the time series value as the third column.
```{r}
eustock <- data.frame(EuStockMarkets)
eustock$time <- time(EuStockMarkets)
df <- eustock |> pivot_longer(1:4, names_to = "Stock")
head(df)
```
Now we use the `import_ts` function included in `tna` to import the data. We indicate the column that identifies each individual time series (`Stock` in our case), the column that contains the numerical value (`value` in our case), and the column that contains the time or order of the time series (`time` for use). Then we need to specify how many discrete states we want to consider.
```{r}
data_ts <- import_ts(df, "Stock", "value", "time", n_states = 4)
```

Now we can plot our time series together with the assigned states at each time point by using the `plot` function and passing the result of `import_ts` as an argument.
```{r}
plot(data_ts)
```
The plot is created with `ggplot2` and uses `facet_wrap`, so you can specify the number of columns you want in your plot. For example, if you want one plot per line, you can pass `ncol = 1`.

```{r}
plot(data_ts, ncol = 1)
```
You can also specify a specific time series by indicating its name as a second argument of the `plot` function.
```{r}
plot(data_ts, "CAC")
```
From `data_ts` we can also create a `tna` model and visualize it, as well as use all of the rest of the features of `tna`.
```{r}
tna_model <- tna(data_ts)
plot(tna_model, cut = 0.01, minimum = 0.005)
```

